### 2015-03-15 staticanalysis.tex
text about the Whoop parsing phase

The proof sketch behind our static lockset analysis is given in Section~\ref{proof}. The verification technique is described in more detail in Section~\ref{method}. The summarisation that we perform to achieve scalability is discussed in Section~\ref{summarisation}. Various optimisations that we perform to achieve precision are described in Section~\ref{optimisations}.

\subsection{Static Lockset Analysis Proof Sketch}
\label{proof}

@@

\subsection{Instrumentation and Analysis}
\label{method}

\whoop Engine is invoked. In this phase, \whoop instruments the program for \emph{static pairwise lockset analysis}

@@

\subsection{Watchdog Summarisation}
\label{summarisation}

@@

\subsection{Optimisations}
\label{optimisations}

@@

\subsection{Assumptions and Limitations}
\label{limitations}

>>>>

The proof sketch behind our static lockset analysis is given in Section~\ref{tech:proof}. The instrumentation and analysis is described in more detail in Section~\ref{tech:method}. The summarisation that we perform to achieve scalability is discussed in Section~\ref{tech:summarisation}. Various optimisations that we perform to achieve precision are described in Section~\ref{tech:optimisations}. Finally, assumption and limitations are discussed in Section~\ref{tech:limitations}.

\subsection{Static Lockset Analysis Proof Sketch}
\label{tech:proof}

@@

\subsection{Instrumentation and Analysis}
\label{tech:method}

The Engine component of \whoop (see Figure~\ref{fig:whoop}) is responsible for parsing, instrumenting and sequentialising the abstract Boogie program. During the initial parsing, \whoop performs a static analysis (on the Boogie IVL code) to identify all lock identifiers and rewrite them to a unique constant Boogie variable, e.g. the following call to \texttt{mutex\_lock()}:

\begin{boogie}
call mutex_lock($p3);
\end{boogie}

Will be rewritten to:

\begin{boogie}
const {:lock} unique lock$0: int;

call mutex_lock(lock$0);
\end{boogie}

If \whoop cannot infer a lock it will exit with a warning. A limitation of lockset analysis is that it cannot understand external locking mechanisms (\whoop currently only supports Linux kernel mutexes and spinlocks). It is relatively straightforward to enhance \whoop, though, with new locking primitives. \whoop will also exit with a warning if it detects ``improper'' use of locks (e.g. dynamic lock creation). The reason behind this is twofold: first, it is arguably infeasible to detect such locks using static analysis; and second, because we want to advocate the use of good locking practises when developing modules for the Linux kernel.

When the lock rewriting finishes, \whoop traverses the Boogie IVL code, uses the information extracted by Chauffeur and separates each independent entry point call-graph from the rest of the program. To achieve this, it duplicates helper functions and renames them accordingly: e.g. for entry point \texttt{ep1}, a helper function \texttt{foo} will be renamed to \texttt{foo\$ep1}. This allows \whoop to perform entry point sensitive instrumentation and analysis. Although the program can potentially become much larger, this does not affect the analysis as discussed later.

%\begin{boogie}[caption = Translation to Boogie IVL using SMACK, label = fig:smack_translation]
%call mutex_lock($p3);
%\end{boogie}

@@

\subsection{Watchdog Summarisation}
\label{tech:summarisation}

@@

\subsection{Optimisations}
\label{tech:optimisations}

@@

\subsection{Assumptions and Limitations}
\label{tech:limitations}
